Empirical Study of Scientific Computer Software: Objectives and Research Questions

### 2020-06-10 OverallResearchProposal/ObjectivesAndResearchQuestions.tex
Restoring previous version of file that was removed by a recent merge

OLD \textbf{What is needed knowledge needs to be captured and what principles, processes and
  methodologies followed to produce sustainable \{software + artifacts\} with
  a reasonable amount of time and energy?}\\

@@

To understand this research objective, we will decompose each constituent part
in turn, starting with the end of the sentence.  In the above objective,
software consists of either programs, which run (execute) on a computer, or
libraries, which provide services to be used by programs.  Both programs and
libraries are created using computer code, written in a programming language.
Many developers place most, if not all, of their emphasis on the computer code
that makes up programs and libraries.  However, we consider the outcome of the
development process to be more than just software and the associated computer
code; we consider the products of the development process to be \{software +
artifacts\}.  The artifacts include code, but also all the other potential work
products, such as documentation, build scripts, test cases, contributor's
guides, etc.  From this point onward, we will use the term \emph{softifacts}
when we are referencing the combination of software and the associated
artifacts.  As we will see, assessing the quality of sustainability is not
possible by only considering the computer code; it depends on all of the
artifacts, and on the processes and methodologies used during software
development.  \wss{Can we consider software as both the computer code and the
  executable (program or library), or should the term software just mean the
  executable?  That is, when we mean code, should we be careful to not instead
  say software?}

Define sustainable.  We start with the general definition:

Might be better to use the adopted definition - since it uses qualities, which
are terms that we need.

What is needed - In the statement of the objective, principles, processes and
methodologies are listed, but not tools.  Tools are needed to support
methodologies, but tools change rapidly.  Our goal is to be more general.  If
the right principles, processes and methodologies are found, the tools will
adapt to match.  However, understanding the current state of the practice will
involve looking at tools, as shown in the later list of research questions.

The notion of ``reasonable amount of time and energy'' is part of the objective,
but reasonable is not defined in the objective.  What is reasonable is going to
depend on the context.  Problems that are large, complex and/or critical are
going to have a greater ``budget'' of time and energy.  When comparing different
alternatives for processes and methodologies, preference would be given to the
alternative that costs less time and energy, assuming that this alternative
achieves the same (or a higher) level of sustainability as its competition.  The
assumptions listed in the next section outline the typical space of problems
where it makes sense to aim for sustainable softifacts.

The objective is not phrased to target a specific software domain, but one of
the research assumptions is that we will focus on Scientific Computing Software
(SCS) examples and applications.  This focus is because creating sustainable
softifacts is more feasible when the software domain is well understood,
which is the case for SCS.  Moreover, SCS is an important, often neglected,
domain of software.

@@

\subsection*{Context, Definitions and Assumptions}

@@

  \emph{Sustainable software means softifacts that are correct,
    reliable, and usable (for the present) and maintainable, portable, reusable
    and reproducible (for the future).}

@@

\subsubsection*{Current State of the Practice Research Questions}

@@

- what is needed to produce softifacts sustainably?

>>>>

@@

To fully explain this research objective, we will decompose each constituent
part in turn, starting with the end of the sentence.  In the above objective,
software consists of either programs, which run (execute) on a computer, or
libraries, which provide services to be used by programs.  Both programs and
libraries are created using computer code, written in a programming language.
Many developers place most, if not all, of their emphasis on the computer code
that makes up programs and libraries.  However, we consider the outcome of the
development process to be more than just software and the associated computer
code; we consider the products of the development process to be \{software +
artifacts\}.  The artifacts include code, but also all the other potential work
products, such as documentation, build scripts, test cases, contributor's
guides, etc.  From this point onward, we will use the term \emph{softifacts}
when we are referencing the combination of software and the associated
artifacts.  As we will explain below, assessing the quality of sustainability is not
possible by only considering the computer code; it depends on all of the
artifacts, and on the processes and methodologies used during software
development.  \wss{Can we consider software as both the computer code and the
  executable (program or library), or should the term software just mean the
  executable?  That is, when we mean code, should we be careful to not instead
  say software?}

We define sustainable softifacts, as follows: \emph{Sustainable softifacts
  satisfy, for a reasonable amount of energy, the software requirements for the
  present, while also being maintainable, reusable and reproducible for the
  future.}  This definition is a specialization of the general definition of
sustainable development from \citet{Brundtland1987}, as discussed in
Section~\ref{Sec_Context}.  The notion of ``reasonable effort'' is part of the
definition of sustainability, but reasonable cannot be completely defined
because what is reasonable is going to depend on the context.  Problems that are
large, complex and/or critical are going to have a greater ``budget'' of time
and energy.  When comparing different alternatives for processes and
methodologies, preference would be given to the alternative that costs less
effort, assuming that this alternative achieves the same (or a higher) level of
sustainability as its competition.  The assumptions listed in the next section
(Section~\ref{Sec_Context}) outline the typical space of problems where it makes
sense to aim for sustainable softifacts.

The overall objectives asks what is needed for sustainability.  This ``what''
can be decomposed further.  We consider the ``what'' to be the principles,
processes and methodologies used for software development.  We exclude tools
from this list because, although tools are needed to support processes and
methodologies, tools change rapidly.  Our goal is to be general.  If the right
principles, processes and methodologies are found, the tools will adapt to
match.  However, understanding the current state of the practice will involve
looking at tools, as shown in the later list of research questions
(Section~\ref{Sec_ResearchQuestions}).

The objective is not phrased to target a specific software domain, but one of
the research assumptions (Section~\ref{Sec_Context}) is that we will focus on
Research Software (RS) examples and applications.  This focus is because
creating sustainable softifacts is more feasible when the software domain is
well understood, which is the case for RS.  Moreover, RS is an important,
often neglected, domain of software.

@@

\subsection*{Context, Definitions and Assumptions} \label{Sec_Context}

@@

  \emph{Sustainable softifacts satisfy, for a reasonable amount of energy, the
    software requirements for the present, while also being maintainable,
    reusable and reproducible for the future.}

@@

\subsubsection*{Current State of the Practice Research Questions} \label{Sec_ResearchQuestions}

@@
